---
title: "8. ANOVA Extensions"
author: "jsg"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y %H:%M')`"
output:
  html_document:
    toc: true
    toc_float: true
    keep_md: true
    self_contained: true
---

Before doing this, review the **8. 2-way Anova** lecture set slides from 
https://sites.google.com/view/biostats/lessons/2-way-anovas-and-interactions and
the **8_Blocking_and_interactions.R** script in the lecture files folder of the
[CUNY-BioStats github repository](https://github.com/jsgosnell/CUNY-BioStats).
Make sure you are comfortable with null and alternative hypotheses and appropriate plots
for all examples.



Remember you should

* add code chunks by clicking the *Insert Chunk* button on the toolbar or by
pressing *Ctrl+Alt+I* to answer the questions!
* **knit** your file to produce a markdown version that you can see!
* save your work often 
  * **commit** it via git!
  * **push** updates to github
  
## Examples

### If interaction is significant

Following the memory example from class, read in and check data

```{r}
memory <- read.table("http://www.statsci.org/data/general/eysenck.txt", header = T,
                     stringsAsFactors = T)
str(memory)
```

Let's put younger level first

```{r}
library(plyr)
memory$Age <- relevel(memory$Age, "Younger")
```

and graph

```{r}
library(Rmisc)
function_output <- summarySE(memory, measurevar="Words", groupvars =
                               c("Age", "Process"), na.rm = T)
library(ggplot2)
ggplot(function_output, aes(x=Age, y=Words,color=Process, 
                                   shape = Process)) +
  geom_line(aes(group=Process, linetype = Process), size=2) +
    geom_point(size = 5) +
  ylab("Words remembered")+ 
  xlab("Age") + 
  ggtitle("Process type interacts with age to impact memory")+
  theme(axis.title.x = element_text(face="bold", size=28), 
        axis.title.y = element_text(face="bold", size=28), 
        axis.text.y  = element_text(size=20),
        axis.text.x  = element_text(size=20), 
        legend.text =element_text(size=20),
        legend.title = element_text(size=20, face="bold"),
        plot.title = element_text(hjust = 0.5, face="bold", size=32))
```

There appears to be some interactions. Let' build a model

```{r}
memory_interactions <- lm(Words ~ Age * Process, memory)
```

and check assumptions.

```{r}
par(mfrow=c(2,2))
plot(memory_interactions)
```

These appear to be met, so look at output

```{r}
library(car)
Anova(memory_interactions, type = "III")
```

Since interaction is significant, analyze subsets. For example, 

```{r}
memory_interactions_young <- lm(Words ~ Process, memory[memory$Age == "Younger",])
plot(memory_interactions_young)
Anova(memory_interactions_young, type = "III")
```

There is a significant difference in words recalled based on process, but why? 
Investigate with post-hoc tests.

```{r}
library(multcomp)
comp_young <- glht(memory_interactions_young, linfct = mcp(Process = "Tukey"))
summary(comp_young)
```

### Blocking example


## Practice

### 1

1. A survey was conducted to see if athletes and non-athletes deal with anger in
the same way.  Data is @

angry <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSaawG37o1ZUEs1B4keIJpZAY2c5tuljf29dWnzqQ0tHNCzfbz85AlWobYzBQ3nPPXJBLP-FWe4BNZB/pub?gid=1784556512&single=true&output=csv", stringsAsFactors = T)

and more information is at 

http://onlinestatbook.com/case_studies/angry_moods.html.

Focus on the following variables:

Sports
1 = athletes, 2 = non-athletes
Gender
1 = males, 2 = females
Expression (AE)
index of general anger expression: 
(Anger-Out) + (Anger-In) - (Control-Out) - (Control-In) + 48

Is there any evidence that gender or athlete status impact how anger is expressed?

### 2

2. A professor carried out a long-term study to see how various factors impacted
pulse rate before and after exercise.  Data can be found at 
http://www.statsci.org/data/oz/ms212.txt
With more info at 
http://www.statsci.org/data/oz/ms212.html.
Is there evidence that frequency of exercise
(Exercise column) and gender impact change in pulse rate for students who ran 
(Ran column = 1)?

### 3

3. (Continuing from question 2) I’m curious if the treatment (running vs not) even had an impact, and if so did it depend on gender. How would you test this?  

