---
title: "4. Continuous tests for 1 population"
author: "jsg"
date: "9/8/2020"
output: github_document
---
Before doing this, review the **Week 3** lecture set slides from 
https://sites.google.com/view/biostats/bio-7800178002/week-3 and
the  **Week_3.R** script in the grad_files_2020 folder of the
[CUNY-BioStats github repository](https://github.com/jsgosnell/CUNY-BioStats).
You’ll also use the qqplot_example.R and
central_limit_theorem.R script from the 
[code_examples folder](https://github.com/jsgosnell/CUNY-BioStats/tree/master/code_examples).

Remember you should

* add code chunks by clicking the *Insert Chunk* button on the toolbar or by
pressing *Ctrl+Alt+I* to answer the questions!
* **knit** your file to produce a markdown version that you can see!
* save your work often 
  * **commit** it via git!
  * **push** updates to github

## Recognizing and assessing normality 

1. Using the qqplot_example.R code, examine the following distributions and, for 
the continuous distributions (marked with a “*”), observe how a normal probability
plot (qqplot) can be used to visually test for approximate normality.

* *Normal (u= 0; σ^2^= 1, 10, 100) 
* *Student's t (df = 1, 10, 30, & 100) 
* *Chi-square (df= 1, 2, 5, 30, 50)
* Bernoulli (P=0.1, 0.5, & 0.9) 
* Binomial (P=0.05; N= 2, 5, 25, & 50); (P=0.25; N= 2, 5, 25, & 50); (P=0.50; 
N= 2, 5, 25, & 50); (P=0.75; N= 2, 5, 25, & 50); (P=0.95; N= 2, 5, 25, & 50) 
* Poisson ( u= 2, 5, 10, 30, & 50)

For this question, its easiest to just source the main file and see what happens.
When you source a script, it is run in R without showing any console output
(but graphs and objects are still produced!).  Try *source("https://raw.githubusercontent.com/jsgosnell/CUNY-BioStats/master/code_examples/qqplot_example.R)*

2. Review the central_limit_theorem.R code if you need to convince/remind 
yourself how common normality of means is for even non-normal data. You can source the code
using the same approach noted in Question 1.

## Just for practice (note some sample sizes may be too small for these to all be good ideas!)

3. Seven observers were shown, for a brief period, a grill with 161 flies impaled 
and were asked to estimate the number. The results are given by Cochran (1954). 
Based on five estimates, they were 183.2, 149.0, 154.0, 167.2, 187.2, 158.0, and 143.0. Test the null hypothesis that the mean of the estimates is 161 flies.

* Assuming variance = 275
* Estimating the variance from the data 
* Using rank transform analysis 
* Using binary transform analysis

4. Yields of 10 strawberry plants in a uniformity trial are given by Baker and 
Baker (1953) as 239, 176, 235, 217, 234, 216, 318, 190, 181, and 225 g. 
Test the hypothesis that µ = 205
* Assuming variance = 1500
* Estimating the variance from the data 
* Using rank transform analysis 
* Using binary transform analysis

4. Evolutionary geneticists predicts the family sex ratio will be 80% female in 
broods of eagles that successfully fledge >3 young. Nests that fledge 3 or more
chicks are very rare but a sample of 30 chicks are obtained from such nests and
they yield 25 females and 5 males. Test the hypotheses that that: 
* a) the sex ratio is 50% females
* b) the sex ratio is 80% females.

5. Studies of flying snakes have led researchers to posit the mean undulation 
rate is 1.4 Hz.  You wish to test this hypothesis using the small sample of 
undulation rates shown below.  Create a small dataset of the paradise tree snake 
undulation rates and choose and justify a test you can use to assess the data.

Undulation rates (in Hz):  0.9, 1.4, 1.2, 1.2, 1.3, 2.0, 1.4, 1.6

6.  Using data from Australian athletes (http://www.statsci.org/data/oz/ais.html 
for details), determine if the average male training at the Australian Institute 
of Sport differs in weight from the average Australian male (85.9 kg) using
bootstrapping techniques. Data at 
```{r}
sport <- read.table("http://www.statsci.org/data/oz/ais.txt", header = T, 
                    stringsAsFactors = T)
```

You can source the bootstrapjsg function using

```{r}
source("https://raw.githubusercontent.com/jsgosnell/CUNY-BioStats/master/code_examples/bootstrapjsg.R")
```

7. Use a Bayesian approach to compare the weight of Australian athletes to 85.9 kg.
Compare approaches that give the null more and less weight.

